---
title: "Tito on the map"
author: "Giorgio Comai"
date: "23 July 2017"
output:   
  html_document:
    code_folding: hide
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggmap, scales, knitr, printr)

```

## Preliminary remarks

This document should be considered draft until it is thoroughly verified. It currently includes substantial inaccuracies.

As of 2017-07-23:

- In the current draft, many data points are repeated multiple times. This [overlay on Google Maps](https://www.google.com/maps/d/viewer?mid=1o575c8xb_TncaQ4Rk8NIxIyNpfY&ll=43.86474657972989%2C19.327976068483963&z=7) shows clearly the issue with the current dataset.

- This means that the absolute number of streets given below is effectively meaningless, and reflects almost exclusively quirkiness of OpenStreetMaps data. Since sometimes the same street is recorded with slightly different names, with slightly different coordinates, the easiest way out of it is probably to keep only one data point per town (even if this would remove, for example, Titov Most in a city where another street dedicated to Tito is present). Other options may be considered

- In the meantime, when many dots are almost one over the other, they should mostly be considered as a single one. Overall, even a polished map would look quite similar. 

- On the other hand, some streets may be missing: they may not be included in OpenStreetMaps, or I may have filtered them out my mistake. The data do not show any self-evident bias, but this does not mean they are correct. 

- Mechanically querying Google Maps for each town in the region may offer more accurate results. 

- Other ideas are welcome. Feel free to open an issue on the repository hosting this file [https://github.com/giocomai/TitoOnTheMap/issues](https://github.com/giocomai/TitoOnTheMap/issues), or to get in touch via [Twitter](http://twitter.com/giocomai/).

## Getting OpenStreetMaps packages

Install [Osmconvert](http://wiki.openstreetmap.org/wiki/Osmconvert) and [Osmfilter](http://wiki.openstreetmap.org/wiki/Osmfilter)

```{r eval=FALSE}
# osmosis not needed? (Osmosis)[http://wiki.openstreetmap.org/wiki/Osmosis/Installation] 
# following code works on linux, for other OS see:
#  http://wiki.openstreetmap.org/wiki/Osmosis/Installation
dir.create(path = "osmosis", showWarnings = FALSE)

if (file.exists(file.path("osmosis", "osmosis-latest.tgz"))==FALSE) {
  download.file(url = "http://bretth.dev.openstreetmap.org/osmosis-build/osmosis-latest.tgz", destfile = file.path("osmosis", "osmosis-latest.tgz"))
  untar(tarfile = file.path("osmosis", "osmosis-latest.tgz"), exdir = "osmosis")
  system(command = "chmod a+x osmosis/bin/osmosis")
}
```

```{r}
# install osmfilter
if (file.exists("osmfilter")==FALSE) {
  system(command = "wget -O - http://m.m.i24.cc/osmfilter.c |cc -x c - -O3 -o osmfilter")
}

# install osmconvert
if (file.exists(file.path("osmconvert64"))==FALSE) {
  download.file(url = "http://m.m.i24.cc/osmconvert64", destfile = file.path("osmconvert64"))
  system(command = "chmod +x osmconvert64")
}


```

## Getting the data from OpenStreetMaps

Download all data for selected countries from [available repositories](http://download.geofabrik.de/europe.html).

```{r}
dir.create(path = "data", showWarnings = FALSE)
dir.create(path = file.path("data", "pbf"), showWarnings = FALSE)
countries <- c("slovenia", "croatia", "bosnia-herzegovina", "serbia", "montenegro", "kosovo", "macedonia")

for (i in countries) {
  if (file.exists(file.path("data", "pbf", paste0(i, "-latest.osm.pbf")))==FALSE) {
    download.file(url = paste0("http://download.geofabrik.de/europe/", i, "-latest.osm.pbf"), destfile = file.path("data", "pbf", paste0(i, "-latest.osm.pbf")))
  }
}

```

Filter only street names, and export them as CSV. (if anyone's familiar with OSM, this is the place to check)

```{r}
# convert to a format that can be read by osmfilter, and remove author data to reduce file size
dir.create(path = file.path("data", "o5m"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "o5m", paste0(i, "-latest.o5m")))==FALSE) {
    system(paste0('./osmconvert64 data/pbf/', i, '-latest.osm.pbf --drop-version --out-o5m -o=data/o5m/', i, '-latest.o5m'))
  }
}

# filter only streets
dir.create(path = file.path("data", "o5m-streets"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "o5m-streets", paste0(i, "-streets.o5m")))==FALSE) {
    system(paste0('./osmfilter data/o5m/', i, '-latest.o5m --keep="highway=*" --drop-version > ', 'data/o5m-streets/', i, '-streets.o5m'))
  }
}

# export to csv only street type, name, and lon/lat

dir.create(path = file.path("data", "csv-streets"), showWarnings = FALSE)

for (i in countries) {
  if (file.exists(file.path("data", "csv-streets", paste0(i, "-streets.csv")))==FALSE) {
    system(paste0('./osmconvert64 data/o5m-streets/', i, '-streets.o5m --all-to-nodes --csv="@id @lat @lon highway name" > data/csv-streets/', i, '-streets.csv', " --csv-separator='; '"))
  }
}

```

## Keep only street names including reference to Tito, and presenting preliminary results

Criteria for filters:

- ends with 'Tito'
- ends with 'Tita'
- includes 'Titov'
- includes 'Тито' (and accordingly, also all 'Титов', etc.)

```{r message=FALSE, warning=FALSE}

StreetsSummary <- data_frame(Country = countries, TotalStreets = as.integer(NA), TitoStreets = as.integer(NA))

FindTito <- "Tito$|Tita$|Titov|Титов|Тито"
tito_all <- data_frame()

for (i in countries) {
  # Import from csv
  temp_streets <- read_delim(file = file.path("data", "csv-streets", paste0(i, "-streets.csv")), delim = "; ", col_names = FALSE, locale = locale(decimal_mark = "."), trim_ws = TRUE)
  # Store total number of streets for each country
  StreetsSummary$TotalStreets[StreetsSummary$Country==i] <- nrow(temp_streets)
  # Filter only Tito's streets
  temp_tito <- temp_streets %>% 
    filter(is.na(X5)==FALSE) %>% 
    filter(stringr::str_detect(string = X5, pattern = stringr::regex(pattern = FindTito, ignore_case = TRUE))) %>%
    transmute(lat = X2, lon = X3, streetname = X5)  %>%
    mutate(country = i)
  # Store number of Tito's streets in given country
  StreetsSummary$TitoStreets[StreetsSummary$Country==i] <- nrow(temp_tito)
  # Merge table for each country
  tito_all <- bind_rows(tito_all, temp_tito)
}
# remove duplicate location with exactly same coordinates
tito_all <- tito_all %>% distinct(lon, lat, .keep_all = TRUE)

write_csv(x = tito_all %>% distinct(), path = "tito_coordinates.csv")
```

## Number of streets dedicated to Tito

```{r}
StreetsSummary %>%
  arrange(TitoStreets) %>% 
  mutate(Country = toupper(Country)) %>% 
  mutate(Country = forcats::fct_inorder(Country)) %>% 
  ggplot(aes(x = Country, y = TitoStreets, label = TitoStreets)) +
  geom_col() +
  scale_y_continuous(name = "", limits = c(0, 150)) +
  scale_x_discrete(name = "") +
  geom_text(hjust = -0.1) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Number of streets dedicated to Tito", subtitle = "(based on OpenStreetMaps data)")

```

## Share of streets dedicated to Tito 

```{r}
options(scipen=999)
StreetsSummary %>%
  mutate(TitoShare = TitoStreets/TotalStreets) %>% 
  arrange(TitoShare) %>% 
  mutate(Country = toupper(Country)) %>% 
  mutate(Country = forcats::fct_inorder(Country)) %>% 
  ggplot(aes(x = Country, y = TitoShare, label = TitoShare)) +
  geom_col() +
  scale_y_continuous(name = "", labels = percent) +
  scale_x_discrete(name = "") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Share of streets dedicated to Tito", subtitle = "(based on OpenStreetMaps data)")
```

```{r}
knitr::kable(StreetsSummary %>% arrange(desc(TitoStreets)))
```


## Mark on the map streets dedicated to Tito

```{r message=FALSE, warning=FALSE}
# Setting the coordinates
wb <- c(left = 12, bottom = 40, right = 26, top = 48)

# Preparing the canvas
mapTonerLite <- get_stamenmap(bbox = wb, zoom = 6, maptype = "toner-lite") %>% ggmap()

mapTonerLite + geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
  labs(x = '', y = '') + ggtitle("Streets and squares dedicated to Tito")
  
  ggsave("TitoTonerLight.png")

# Preparing the canvas  
mapG <- get_googlemap("Sarajevo", scale = 2, zoom = 6) %>% ggmap()

mapG + geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
  labs(x = '', y = '') + ggtitle("Streets and squares dedicated to Tito")

ggsave("TitoGmaps.png")


# # Preparing the canvas
# mapWatercolor <- get_stamenmap(bbox = wb, zoom = 7, maptype = "watercolor") %>% ggmap()
# # Adding the dots
# mapWatercolor + geom_point(data=tito_all, aes(x=lon, y=lat), color="red", size=2, alpha=0.5) +
#   labs(x = '', y = '') + ggtitle("Streets dedicated to Tito")
# ggsave("TitoWatercolor.png")

```

## Summary of all street names found

N.B. Check if something is missing and update criteria

```{r}
knitr::kable(tito_all %>% select(country, streetname) %>% group_by(country, streetname) %>% count() %>% arrange(country, desc(n), streetname))
write_csv(x = tito_all %>% select(country, streetname) %>% group_by(country, streetname) %>% count() %>% arrange(country, desc(n), streetname), path = "tito_streetnames.csv")
```

## About the author

Giorgio Comai is on [Twitter](http://twitter.com/giocomai/). His website is [https://giorgiocomai.eu/](https://giorgiocomai.eu/)
